Inference for Categorical Data

### 2018-11-02 DATA606/Inference for Categorical Data.Rmd
More work on Inference for Categorical Data

(a) We are 95% confident that between 43% and 49% of Americans in this sample support the decision of the U.S. Supreme Court on the 2010 healthcare law.

(b) We are 95% confident that between 43% and 49% of Americans support the decision of the U.S. Supreme Court on the 2010 healthcare law.

(c) If we considered many random samples of 1,012 Americans, and we calculated the sample proportions of those who support the decision of the U.S. Supreme Court, 95% of those sample proportions will be between 43% and 49%.

(d) The margin of error at a 90% confidence level would be higher than 3%.

@@

(a) Is 48% a sample statistic or a population parameter? Explain.

(b) Construct a 95% confidence interval for the proportion of US residents who think marijuana should be made legal, and interpret it in the context of the data.

(c) A critic points out that this 95% confidence interval is only accurate if the statistic follows a normal distribution, or if the normal model is a good approximation. Is this true for these data? Explain.

(d) A news piece on this survey’s findings states, “Majority of Americans think marijuana should be legalized.” Based on your confidence interval, is this news piece’s statement justified?

@@

As discussed in Exercise 6.12, the 2010 General Social Survey reported a sample where about 48% of US residents thought marijuana should be made legal. If we wanted to limit the margin of error of a 95% confidence interval to 2%, about how many Americans would we need to survey ?

@@

According to a report on sleep deprivation by the Centers for Disease Control and Prevention, the proportion of California residents who reported insufficient rest or sleep during each of the preceding 30 days is 8.0%, while this proportion is 8.8% for Oregon residents. These data are based on simple random samples of 11,545 California and 4,691 Oregon residents. Calculate a 95% confidence interval for the difference between the proportions of Californians and Oregonians who are sleep deprived and interpret it in context of the data.

@@

(b) What type of test can we use to answer this research question?

(c) Check if the assumptions and conditions required for this test are satisfied.

@@

## 6.48 Coffee and Depression.

@@

(a) What type of test is appropriate for evaluating if there is an association between coffee intake and depression?

(b) Write the hypotheses for the test you identified in part (a).

(c) Calculate the overall proportion of women who do and do not suffer from depression.

(d) Identify the expected count for the highlighted cell, and calculate the contribution of this cell to the test statistic, i.e. $\frac{(\text{Observed} − \text{Expected})^2}{\text{Expected}}$.

@@

(f) What is the conclusion of the hypothesis test?

(g) One of the authors of this study was quoted on the NYTimes as saying it was “too early to recommend that women load up on extra coffee” based on just this study.64 Do you agree with this statement? Explain your reasoning.

>>>>

(a) We are 95% confident that between 43% and 49% of Americans in this sample support the decision of the U.S. Supreme Court on the 2010 healthcare law.**False.  The CI applies to the population not the sample.**

(b) We are 95% confident that between 43% and 49% of Americans support the decision of the U.S. Supreme Court on the 2010 healthcare law.**True.  We are confident the population parameter falls within the range.**

(c) If we considered many random samples of 1,012 Americans, and we calculated the sample proportions of those who support the decision of the U.S. Supreme Court, 95% of those sample proportions will be between 43% and 49%.**True.  95% of the area under the curve would be between 43% and 49%.**

(d) The margin of error at a 90% confidence level would be higher than 3%.**False.  A higher confidence level would have a larger margin of error.**

@@

(a) Is 48% a sample statistic or a population parameter? Explain. **Sample statistics because it is the share of the respondents.**

(b) Construct a 95% confidence interval for the proportion of US residents who think marijuana should be made legal, and interpret it in the context of the data.
```{r}
n <- 1259
p <- 0.48
z <- 1.96
se <- sqrt((p*(1-p))/n)
me <- z * se
```
**The 95% confidence interval is (`r p - me`, `r p + me`).  That means that the between 45% and 51% of US residents thing marijuana should be made legal.**

(c) A critic points out that this 95% confidence interval is only accurate if the statistic follows a normal distribution, or if the normal model is a good approximation. Is this true for these data? Explain.**There are `r round(p * 1259, 0)` people who agree and `r round((1-p) * 1259, 0)` people who disagree.  This is larger than the 10 success/failure rule of thumb.  The observations are probably independent and the sample size is about right to represent the United States.  The critics points are not valid.**

(d) A news piece on this survey’s findings states, “Majority of Americans think marijuana should be legalized.” Based on your confidence interval, is this news piece’s statement justified? **No.  This is an exageration.**

@@

As discussed in Exercise 6.12, the 2010 General Social Survey reported a sample where about 48% of US residents thought marijuana should be made legal. If we wanted to limit the margin of error of a 95% confidence interval to 2%, about how many Americans would we need to survey ? $0.02 = 1.96 \times \sqrt{\frac{0.5 \times 0.5}{n}} \rightarrow 0.02^2 = 1.96^2 \times \frac{0.25}{n} \rightarrow 0.0004 = \frac{3.8416 \times 0.25}{n} \rightarrow n = \frac{0.9604}{0.0004} \rightarrow n = 2401$ **You would need to survey 2,401 people.**

@@

According to a report on sleep deprivation by the Centers for Disease Control and Prevention, the proportion of California residents who reported insufficient rest or sleep during each of the preceding 30 days is 8.0%, while this proportion is 8.8% for Oregon residents. These data are based on simple random samples of 11,545 California and 4,691 Oregon residents. Calculate a 95% confidence interval for the difference between the proportions of Californians and Oregonians who are sleep deprived and interpret it in context of the data.
```{r}
p1 <- 0.08
p2 <- 0.088
n1 <- 11545
n2 <- 4691
z <- 1.96
se <- sqrt(((p1*(1-p1))/n1)+((p2*(1-p2))/n2))
me <- z * se
```
**The 95% confidence interval is (`r (p1-p2)-me`, `r (p1-p2)+me`).  This means the difference between CA and OR is between -2% and 0%.**

@@

**H$_0$: The proportion between the groups are the same: $p_1$ = $p_2$ = $p_3$ = $p_4$.**

**H$_A$: At least one of the proportions are different from the others.**

(b) What type of test can we use to answer this research question? **Chi-squared Test or goodness of fit test.**

(c) Check if the assumptions and conditions required for this test are satisfied. **Independence: each case only contributes to one cell of the table and are a independent. Sample size: Each cell of expected cases should have at least 5 cases.  We'll have to see if this holds.**

@@

|          | Woods | Cultivated grassplot | Deciduous forests | Other | Total |
|----------|-------|----------------------|-------------------|-------|-------|
| Observed | 4     | 16                   | 61                | 345   | 426   |
| Expected | 20    | 63                   | 169               | 174   | 426   |

```{r}
woods <- (4 - 20)^2 / 20
cult <- (16 - 63)^2 / 63
deci <- (61 - 169)^2 / 169
other <- (345 - 174)^2 / 174
k <- 4
chi_sq <- woods + cult + deci + other
df <- k - 1
pchisq(chi_sq, df, lower.tail = FALSE)
```
**The P value is smaller than 0.05 so the evidence suggest we reject the null hypothesis, meaning that the deer do not prefer to forage in all types of terain equally.**

## 6.48 Coffee and Depression

@@

(a) What type of test is appropriate for evaluating if there is an association between coffee intake and depression? **A Chi-squared or goodness of fit test.**

(b) Write the hypotheses for the test you identified in part (a).
**H$_0$: The proportion between the groups are the same: $p_1$ = $p_2$ = $p_3$ = $p_4$ = $p_5.**

**H$_A$: At least one of the proportions are different from the others.**

(c) Calculate the overall proportion of women who do and do not suffer from depression. **`r 48132/50739`**

(d) Identify the expected count for the highlighted cell, and calculate the contribution of this cell to the test statistic, i.e. $\frac{(\text{Observed} − \text{Expected})^2}{\text{Expected}}$

**Expected is defined as:** $2607 \times
\frac{6617}{50739} = 2607 \times 0.1304125 = 340$

**so $\frac{(373 − 340)^2}{340} = \frac{1089}{340} = 3.203$.**

@@

```{r}
k <- 5
chi_sq <- 20.93
df <- k - 1
pchisq(chi_sq, df, lower.tail = FALSE)
```

(f) What is the conclusion of the hypothesis test? **The p-value is lower than 0.05 therefore you would reject the null hypothesis.**

(g) One of the authors of this study was quoted on the NYTimes as saying it was “too early to recommend that women load up on extra coffee” based on just this study.64 Do you agree with this statement? Explain your reasoning.**Yes.  It's just one study.  There is a chance that it is a fluke and should be confirmed with more research.**
